Positivity-Preserving Adaptive Runge--Kutta Methods

### 2020-03-31 nusslein_vsrp/latex/overview.tex

Our means to ensure positivity can be interpreted as a projection
approach, where the numerical solution is adapted to satisfy the
positivity constraint at the end of each step. In contrast to the
usual orthogonal projection, which has also been proposed to deal
with positivity constraints \cite{shampine1986conservation},
our approach preserves all linear invariants of the given ODE. These
invariants can be very important. e.g.\ the total mass for a transport
problem or in reaction systems. The property to preserve all linear
invariants is shared by RRK methods and has been shown to be an
important advantage compared to orthogonal projection methods
\cite{ranocha2020relaxationHamiltonian}.

@@

>>>>

Our means to ensure positivity can be interpreted as a projection
approach, where the numerical solution is adapted to satisfy the
positivity constraint at the end of each step. In contrast to the
usual orthogonal projection, which has also been proposed to deal
with positivity constraints \cite{shampine1986conservation},
our approach preserves all linear invariants of the given ODE. These
invariants can be very important. e.g.\ the total mass for a transport
problem or in reaction systems. The property to preserve all linear
invariants is shared by RRK methods and has been shown to be an
important advantage compared to orthogonal projection methods
\cite{ranocha2020relaxationHamiltonian}. Of course, it is also
possible to enforce the preservation of linear invariants in projection
methods, but these invariants have to be known explicitly
\cite{sandu2001positive}.

@@

In contrast to other projection methods
\cite{shampine1986conservation,sandu2001positive},
minimizing the deviation of the weights instead of the deviation
of the projected solution is computationally much more efficient
for large systems, arising for example in the discretization of
PDEs.
